Building a GNU/Hurd Cluster

### 2018-01-26 slides/rpctrace.tex
rpctrace slide series - add slide series illustrating insert_right problem

\usetikzlibrary{positioning, calc}

@@

\tikzstyle{send} = [rectangle, fill=green, rounded corners] 
\tikzstyle{recv} = [rectangle, fill=red!50, rounded corners]
\tikzstyle{empty} = [minimum size=0, inner sep=0]
\tikzstyle{queue} = [draw, rectangle, minimum height=.5cm, minimum width=1cm, pattern=vertical lines]

@@

\only<1,8->{\node (R1) at (PUT Port 0) [recv, anchor=west] {R};}
\only<2-7>{\node (R1) at (PUT Port 0) [recv, anchor=west, minimum size=0, inner sep=0] {};}
\node (r1) at (R1) {};

\only<1-3,7->{
  \node (S1) at (R1.east) [send, anchor=west] {S};
  \node (S2) at (S1.east) [send, anchor=west] {S};
}

@@

\only<3-9>{
  \node (S3) at (R2.east) [send, anchor=west] {S};
}
\only<10->{
  \node (S3) at (R2.east) [empty] {};
}

\only<5-7>{
  \node (R3) at (rpctrace Right Port 0) [recv, anchor=east] {R};
}
\only<8->{
  \node (R3) at (rpctrace Right Port 0) [minimum size=0, inner sep=0] {};
}

\only<5->{
  \draw [violet, thick] (rpctrace Right Port 0) -| ($(rpctrace Right Port 0)!.5!(queue2.west)$) |- (queue2.west);
}

\only<6->{
  \node (S4) at (R3.west) [send, anchor=east] {S};
}

\only<7->{
  \draw [violet, thick] (queue2.east) -| ($(queue2.east)!.5!(PUT Port 0)$) |- (PUT Port 0);
}

\only<9->{
  % S3's location moves between slides, so starting this line at S4 keeps the dashes from morphing between slides
  \draw [violet, dashed, thick] (S4) -- (S3);
}

>>>>

\usetikzlibrary{positioning, calc, arrows, fit}

@@

\tikzstyle{send} = [rectangle, fill=green, rounded corners]
\tikzstyle{recv} = [rectangle, fill=red!50, rounded corners]
\tikzstyle{empty} = [minimum size=0, inner sep=0]
\tikzstyle{queue} = [draw, rectangle, minimum height=.5cm, minimum width=1cm, pattern=vertical lines]
\tikzstyle{message} = [draw, fill=red!20, text width=5em, text depth=3em, text centered,
                       minimum width=5em, minimum height=2em, rounded corners]

\tikzstyle{port link} = [black, -triangle 90]
\tikzstyle{message path} = [violet, thick, -triangle 90]
\tikzstyle{right link} = [red, thick, dotted, ->]
\tikzstyle{internal link} = [dashed, thick]

% illustration of rpctrace problems with mach_port_insert_right()

\begin{frame}[fragile]
\begin{block}{}
\begin{semiverbatim}
\tiny
ports\_create\_port (port\_class, port\_bucket, sizeof (struct port\_info), \&readport);
readportname = ports\_get\_right (readport);
mach\_port\_insert\_right (mach\_task\_self (), readportname, readportname, MACH\_MSG\_TYPE\_MAKE\_SEND);
\end{semiverbatim}
\end{block}

\begin{block}{}
\begin{semiverbatim}
\tiny
routine mach\_port\_insert\_right(
                task            : ipc\_space\_t;
                name            : mach\_port\_name\_t;
                poly            : mach\_port\_poly\_t);

\end{semiverbatim}
\end{block}

\vfill

\begin{center}
\begin{tikzpicture}

\task [left = of center point] {rpctrace}{}{lr}{
  \only<1-3>{
    \node (rpctrace Message) [message, anchor=north, opacity=0] {\scriptsize insert\_right};
  }
  \only<4-5>{
    \node (rpctrace Message) [message, anchor=north] {\scriptsize insert\_right};
  }
  \only<6>{
    \node (rpctrace Message) [message, anchor=north, fill=none, dashed, text opacity=0] {\scriptsize insert\_right};
  }
}

\only<4-5>{
  \node (rpctrace sendright) at (rpctrace Message) [send, anchor=west] {S};
  \node at (rpctrace sendright.west) [anchor=east, outer sep=0, inner sep=0] {(10,};
  \node at (rpctrace sendright.east) [anchor=west, outer sep=0, inner sep=0] {)};
}

\task [right = of center point] {Traced Task}{}{lk}{
  \only<1-2,5->{
    \node (Traced Task Message) [message, anchor=north, opacity=0] {\scriptsize insert\_right};
  }
  \only<3>{
    \node (Traced Task Message) [message, anchor=north] {\scriptsize insert\_right};
  }
  \only<4>{
    \node (Traced Task Message) [message, anchor=north, fill=none, dashed, text opacity=0] {\scriptsize insert\_right};
  }
}

\only<3>{
  \node (Traced Task sendright) at (Traced Task Message) [send, anchor=west] {S};
  \node at (Traced Task sendright.west) [anchor=east, outer sep=0, inner sep=0] {(10,};
  \node at (Traced Task sendright.east) [anchor=west, outer sep=0, inner sep=0] {)};
}

\draw [port link] (rpctrace Port 7) -| ($(rpctrace Port 7)!.5!(Traced Task Task Port)$) |- (Traced Task Task Port);

\node (S1) at (rpctrace Port 7) [send, anchor=east] {S};
\node (R1) at (rpctrace Port 4) [recv, anchor=east] {R};

\draw [internal link] (S1) -- +(-.5,0) |- (R1);

\draw [port link] (Traced Task Port 4) -- (rpctrace Port 4);
\node (S2) at (Traced Task Port 4) [send, anchor=west] {S};
\node at (S2.east) [anchor=west] {\scriptsize mach\_task\_self()};

\only<4>{
  \draw [message path] (Traced Task Message.north) |- (S2);
  \draw [message path] (Traced Task Port 4) -- (rpctrace Port 4);
  \draw [message path] (R1) -| (rpctrace Message.north);
}

\only<6>{
  \draw [message path] (rpctrace Message.north) |- (S1);
  \draw [message path] (rpctrace Port 7) -| ($(rpctrace Port 7)!.5!(Traced Task Task Port)$) |- (Traced Task Task Port);
  \draw [message path] (Traced Task Task Port) -| ($(Traced Task Task Port)!.5!(Kernel Message.west)$) |- (Kernel Message.west);
}

\only<2->{
  \node (R2) at (Traced Task Port -4) [recv, anchor=west] {R};
  \node at (R2.north) [anchor=south west, inner sep=2] {10};
}

\only<3>{
  \draw [right link] (Traced Task sendright) -- (R2);
}

\only<4>{
  \draw [right link] (rpctrace sendright) -- (R2);
}

\only<5->{
  \node (S3) at (rpctrace Port -4) [send, anchor=east] {S};
  \node (R3) at (rpctrace Port -1) [recv, anchor=east] {R};
  \draw [internal link] (S3) -- +(-.5,0) |- (R3);
  \draw [port link] (rpctrace Port -4) -- (Traced Task Port -4);
}

\only<5>{
  \draw [right link] (rpctrace sendright) -- (R3);
}

\only<1-5>{
  \node (Kernel Message) [message, anchor=south west, opacity=0] at (Traced Task.north) {\scriptsize insert\_right};
}
\only<6->{
  \node (Kernel Message) [message, anchor=south west] at (Traced Task.north) {\scriptsize insert\_right};
  \node (Kernel sendright) at (Kernel Message) [send, anchor=west] {S};
  \node at (Kernel sendright.west) [anchor=east, outer sep=0, inner sep=0] {(10,};
  \node at (Kernel sendright.east) [anchor=west, outer sep=0, inner sep=0] {)};
  \draw [right link] (Kernel sendright) -- (R3);
}

\end{tikzpicture}
\end{center}

\end{frame}

%% illustration of attach operation (two frames)

@@

\only<1,5->{\node (R1) at (PUT Port 0) [recv, anchor=west] {R};}
\only<2-4>{\node (R1) at (PUT Port 0) [empty] {};}

\only<1-3,6->{
  \node (S1) at (R1.east) [send, anchor=west] {S};
  \node (S2) at (S1.east) [send, anchor=west] {S};
}

@@

\only<3-8>{
  \node (S3) at (R2.east) [send, anchor=west] {S};
}
\only<9->{
  \node (S3) at (R2.east) [empty] {};
}

\only<7->{
  \node (S4) at (rpctrace Right Port 0) [send, anchor=east] {S};
  \draw [violet, thick] (rpctrace Right Port 0) -| ($(rpctrace Right Port 0)!.5!(queue2.west)$) |- (queue2.west);
}

\only<8->{
  % S3's location moves between slides, so starting this line at S4 keeps the dashes from morphing between slides
  \draw [violet, dashed, thick] (S4) -- (S3);
}
